# Reproducible Research: Peer Assessment 1

```{r, echo=FALSE, results='hide', warning=FALSE, message=FALSE}

```

## Loading and preprocessing the data
### 1. Load the data (i.e. read.csv())
##### => Read the file and the packages that are needed.
```{r, results='markup', warning=TRUE, message=TRUE}
library(ggplot2)
library(scales)
library(Hmisc)
actdata <- read.csv('activity.csv')
```
### 2. Process/transform the data (if necessary) into a format suitable for your analysis
##### => Transform the "date" column into the type of "Date"
```{r}
actdata$date <- as.Date(actdata$date)
```

-----

## What is mean total number of steps taken per day?
    ```{r}
stepsperday <- tapply(actdata$steps, actdata$date, sum, na.rm=TRUE)
```

### 1. Make a histogram of the total number of steps taken each day
```{r}
qplot(stepsperday, xlab='Total Steps Per Day')
```

### 2. Calculate and report the mean and median total number of steps taken per day
```{r}
stepsperdaymean <- mean(stepsperday)
stepsperdaymedian <- median(stepsperday)
```
* Mean: `r stepsperdaymean`
* Median:  `r stepsperdaymedian`

-----
    
## What is the average daily activity pattern?
    ```{r}
avgsteps <- aggregate(x=list(avgsteps=actdata$steps), by=list(interval=actdata$interval), FUN=mean, na.rm=TRUE)
```

### 1. Make a time series plot
```{r}
ggplot(data=avgsteps, aes(x=interval, y=avgsteps)) +
    geom_line() +
    xlab("5-Minute Interval") +
    ylab("Average Number of Steps Taken") 
```

### 2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
maxinterval <- avgsteps$interval[which(avgsteps$avgsteps == max(avgsteps$avgsteps))]
```

* Most Steps at: `r maxinterval`

----
    
## Imputing missing values
### 1. Calculate and report the total number of missing values in the dataset 
    ```{r}
NumMissing <- length(which(is.na(actdata$steps)))
```

* Number of missing values: `r NumMissing`

### 2. Devise a strategy for filling in all of the missing values in the dataset.
### 3. Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
actfill <- actdata
actfill$steps <- impute(actdata$steps, fun=mean)
```


### 4. Make a histogram of the total number of steps taken each day 
```{r}
stepsbydayfill <- tapply(actfill$steps, actfill$date, sum)
qplot(stepsbydayfill, xlab='Total steps per day (Imputed)', ylab='Frequency using binwith 500', binwidth=500)
```

### ... and Calculate and report the mean and median total number of steps taken per day. 
```{r}
stepsbydayfillmean <- mean(stepsbydayfill)
stepsbydayfillmedian <- median(stepsbydayfill)
```
* Mean (Imputed): `r stepsbydayfillmean`
* Median (Imputed):  `r stepsbydayfillmedian`


----
    
## Are there differences in activity patterns between weekdays and weekends?
##### 1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
    
    ```{r}
actfill$dateType <-  ifelse(as.POSIXlt(actfill$date)$wday %in% c(0,6), 'weekend', 'weekday')
```

### 2. Make a panel plot containing a time series plot

```{r}
avgactfill <- aggregate(steps ~ interval + dateType, data=actfill, mean)
ggplot(avgactfill, aes(interval, steps)) + 
    geom_line() + 
    facet_grid(dateType ~ .) +
    xlab("5-minute interval") + 
    ylab("avarage number of steps")
```